
---
title: "Exercise Set 1"
author: "By Mofrad Muntasir"
output: html_document
---


<br>

The purpose of this exercise is to become familiar with:

1. Basic statistics functions in R;
2. Simple matrix operations;
3. Simple data manipulations; 
4. The idea of functions as well as some useful customized functions provided. 

While doing this exercise we will also see how to generate replicable and customizable reports. For this purpose the exercise uses the R Markdown capabilities (see [Markdown Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf) or a [basic introduction to R Markdown](http://rmarkdown.rstudio.com/authoring_basics.html)).  These capabilities allow us to create dynamic reports. For example today's date is `r Sys.Date()` (you need to see the .Rmd to understand that this is *not* a static typed-in date but it changes every time you compile the .Rmd - if the date changed of course).

Before starting, make sure you have pulled the [exercise files](https://github.com/InseadDataAnalytics/INSEADAnalytics/tree/master/Exercises/Exerciseset1)  on your github repository (if you pull the course github repository you also get the exercise set files automatically). Moreover, make sure you are in the directory of this exercise. Directory paths may be complicated, and sometimes a frustrating source of problems, so it is recommended that you use these R commands to find out your current working directory and, if needed, set it where you have the main files for the specific exercise/project (there are other ways, but for now just be aware of this path issue). For example, assuming we are now in the "MYDIRECTORY/INSEADAnalytics" directory, we can do these: 

```{r echo=TRUE, eval=FALSE, tidy=TRUE}
#getwd()

#setwd("Exercises/Exerciseset1/")

#list.files()
```

**Note:** you can always use the `help` command in Rstudio to find out about any R function (e.g. type `help(list.files)` to learn what the R function `list.files` does).

Let's now see the exercise. 

**IMPORTANT:** You should answer all questions by simply adding your code/answers in this document through editing the file ExerciseSet1.Rmd and then clicking on the "Knit HTML" button in RStudio.  Once done, please post your .Rmd and html files in your github repository. 

<hr>
<hr>

### Exercise Data

We download daily prices (open, high, low, close, and adjusted close) and volume data of publicly traded companies and markets from the web (e.g. Yahoo! or Google, etc). This is done by sourcing the file data.R as well as some helper functions in herpersSet1.R which also installs a number of R libraries (hence the first time you run this code you will see a lot of red color text indicating the *download* and *installation* process):

```{r eval = TRUE, echo=TRUE, error = FALSE, warning=FALSE,message=FALSE,results='asis'}
source("helpersSet1.R")
source("dataSet1.R")
```

For more information on downloading finance data from the internet as well as on finance related R tools see these starting points (there is a lot more of course available):

* [Some finance data loading tools](http://www.r-bloggers.com/r-code-yahoo-finance-data-loading/)
* [Connecting directly to Bloomberg](http://www.r-bloggers.com/rblpapi-connecting-r-to-bloomberg/)
* [Some time series plot tools](http://www.r-bloggers.com/plotting-time-series-in-r-using-yahoo-finance-data/)
* [Various finance code links](https://cran.r-project.org/web/views/Finance.html)
* [More links](http://blog.revolutionanalytics.com/2013/12/quantitative-finance-applications-in-r.html)
* [Even more links](http://www.r-bloggers.com/financial-data-accessible-from-r-part-iv/)
* Of course endless available code (e.g. like this one that seems to [get companies' earnings calendars](https://github.com/gsee/qmao/blob/master/R/getCalendar.R))

#### Optional Question 

1. Can you find some interesting finance related R package or github repository? 

**Your Answers here:**

<br>


<hr>
<hr>

### Part I: Statistics of S&P Daily Returns

We have `r nrow(StockReturns)` days of data, starting from `r rownames(StockReturns)[1]` until `r tail(rownames(StockReturns),1)`.  Here are some basic statistics about the S&P returns:

1. The cumulative returns of the S&P index during this period is `r round(100*sum(StockReturns[,1]),1)`%.
2. The average daily returns of the S&P index during this period is `r round(100*mean(StockReturns[,1]),3)`%;
2. The standard deviation of the daily returns of the S&P index during this period is `r round(100*sd(StockReturns[,1]),3)`%;

Here are returns of the S&P in this period (note the use of the helper function pnl_plot - defined in file helpersSet1.R):

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig.height=4,fig.width= 6, fig=TRUE}
SPY = StockReturns[,"SPY"]
pnl_plot(SPY)
```

#### Questions

1. Notice that the code also downloads the returns of Apple during the same period. Can you explain where this is done in the code (including the .R files used)?
2. What are the cumulative, average daily returns, and the standard deviation of the daily returns of Apple in the same period? 
3. *(Extra points)* What if we want to also see the returns of another company, say Yahoo!, in the same period? Can you get that data and report the statistics for Yahoo!'s stock, too? 

**Your Answers here:**
<br><b>1.</b> Under myticker, (mytickers = c("SPY", "AAPL") ) was used. Which makes the software identify SPY and AAPL (Apple)for SPY and Apple. The following formula then downloads all the stocks mentioned in myticker. 

  for (ticker_index in 1:length(mytickers)){
    ticker_to_get = mytickers[ticker_index]
    print(paste("\nDownloading ticker ", ticker_to_get, " ..."))
    
    The code is written in the file titled "dataset1.R"

<b> 2. </b>
```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig.height=4,fig.width= 6, fig=TRUE}
AAPL = StockReturns[,"AAPL"]
pnl_plot(AAPL)
```
<br> The plot was made using the following code : AAPL = StockReturns[,"AAPL"]
pnl_plot(AAPL)

<br> a. The cumulative returns of the Apple during this period is `r round(100*sum(StockReturns[,"AAPL"]),1)`%.
b. The average daily returns of the Apple during this period is `r round(100*mean(StockReturns[,"AAPL"]),3)`%;
c. The standard deviation of the daily returns of Apple during this period is `r round(100*sd(StockReturns[,"AAPL"]),3)`%;

<br> <b> 3. </b> By adding "YHOO" code in myticker, we can update the dataset to download Yahoo data as well. After that we can run the same formula to get the analysis on Yahoo. 

{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig.height=4,fig.width= 6, fig=TRUE}
YHOO = StockReturns[, "YHOO"]
pnl_plot(YHOO)

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig.height=4,fig.width= 6, fig=TRUE}
YHOO = StockReturns2 [, "YHOO"]
pnl_plot(YHOO)
```
<br> a. The cumulative returns of the Yahoo during this period is 290.6%.
b. The average daily returns of the Yahoo during this period is 0.072`%;
c. The standard deviation of the daily returns of Yahoo during this period is `r 3.019%;


<hr>
<hr>

### Part II: Simple Matrix Manipulations

For this part of the exercise we will do some basic manipulations of the data. First note that the data are in a so-called matrix format. If you run these commands in RStudio (use help to find out what they do) you will see how matrices work: 

```{r eval = FALSE, echo=TRUE}
class(StockReturns)
dim(StockReturns)
nrow(StockReturns)
ncol(StockReturns)
StockReturns[1:4,]
head(StockReturns,5)
tail(StockReturns,5) 
```

We will now use an R function for matrices that is extremely useful for analyzing data. It is called *apply*. Check it out using help in R. 

For example, we can now quickly estimate the average returns of S&P and Apple (of course this can be done manually, too, but what if we had 500 stocks - e.g. a matrix with 500 columns?) and plot the returns of that 50-50 on S&P and Apple portfolio:

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig=TRUE}
portfolio = apply(StockReturns,1,mean)
names(portfolio) <- rownames(StockReturns)
pnl_plot(portfolio)
```


We can also transpose the matrix of returns to create a new "horizontal" matrix. Let's call this matrix (variable name) transposedData. We can do so using this command:  `r transposedData = t(StockReturns)`. 

#### Questions

1. What R commands can you use to get the number of rows and number of columns of the new matrix called transposedData?
2. Based on the help for the R function *apply* (`help(apply)`), can you create again the portfolio of S&P and Apple and plot the returns in a new figure below?

**Your Answers here:**
<br> <b> 1. </b> Tranposed Data (transposedData) is the reorganized but same data from Stock Returns. We get that by doing the following : 
transposedData = t(stockReturns)

For counting the number of rows and columns, we need to use nrow(transposedData) and ncol(transposedData) formula. So, using these, the number of rows and columns are : `r nrow(transposedData)` and `r ncol(transposedData)`
<br> <b> 2 </b> 
<br> In this step, my first step is to use transposedData to change the stockreturns data's format. After that, I have used apply formula to use transposedData to create the same chart. In the formula, the second element dictates whether to use row or column. As I have put 2 there, the software will use the columns. Following is the formula : "portfolio_TP = apply(transposedData,2,mean)
names(portfolio_TP) <- rownames(StockReturns)
pnl_plot(portfolio_TP)

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig=TRUE}
portfolio_TP = apply(transposedData,2,mean)
names(portfolio_TP) <- rownames(StockReturns)
pnl_plot(portfolio_TP)
```


The difference between this method and the earlier is in the second part of apply formula. The second part dictates whether to use row or column from the dataset. As we used row (1) in the previous data, and transposed data just converts row data into column data, I used column (2) in this formula to get to the same result.
<br>

<hr>
<hr>

### Part III: Reproducibility and Customization

This is an important step and will get you to think about the overall process once again. 

#### Questions

1. We want to re-do all this analysis with data since 2001-01-01: what change do we need to make in the code (hint: all you need to change is one line - exactly 1 number! - in data.R file), and how can you get the new exercise set with the data since 2001-01-01? 
2. *(Extra Exercise)*  Can you get the returns of a few companies and plot the returns of an equal weighted portfolio with those companies during some period you select? 

**Your Answers here:**
<br> <b> 1 </b> As 1st January 2001 was a holiday, there's no data for that day. But we can get data from December 29th by changing the start date in dataset1.R to 2000-12-29
<br>
<br> <b> 2 </b> I have added stock returns of Google and Tesla to the mix by creating a new dataset that titled StockReturns2. StockReturns 2 only downloads data from July 11, 2012 to July 10, 2013.



```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis',fig.align='center', fig=TRUE}
portfolio_New = apply(StockReturns2,1,mean)
names(portfolio_New) <- rownames(StockReturns2)
pnl_plot(portfolio_New)
```


<br>

<hr>
<hr>

###  Part IV: Read/Write .CSV files

Finally, one can read and write data in .CSV files. For example, we can save the first 20 days of data for S&P and Apple in a file using the command:

```{r eval = TRUE, echo=TRUE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
write.csv(StockReturns[1:20,c("SPY","AAPL")], file = "twentydays.csv", row.names = TRUE, col.names = TRUE) 
```

Do not get surpsised if you see the csv file in your directories suddenly! You can then read the data from the csv file using the read.csv command. For example, this will load the data from the csv file and save it in a new variable that now is called "myData": 

```{r eval = TRUE, echo=TRUE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
myData <- read.csv(file = "twentydays.csv", header = TRUE, sep=";")
```

Try it!

#### Questions

1. Once you write and read the data as described above, what happens when you run this command in the console of the RStudio: `sum(myData != StockReturns[1:20,])`
2. *(Extra exercise)* What do you think will happen if you now run this command, and why:  

```{r eval = FALSE, echo=TRUE}
myData + StockReturns[1:40,]
```

**Your Answers here:**
<br> <b> 1 </b> 
```{r eval = TRUE, echo=TRUE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
write.csv(StockReturns[1:20,c("SPY","AAPL")], file = "twentydays.csv", row.names = TRUE, col.names = TRUE)

```

`r myData <- read.csv(file = "twentydays.csv", header = TRUE, sep=";")` 
`r sum(myData != StockReturns[1:20,])`

The result of this will be 20.

<br> <b> 2 </b> There's an error "Error in FUN (left, right) : non-numeric argument to binary operator". If we read the data we see that it is not a string of data rather a line with data, comma, dates and semi colon. So there is not a way to get this to work.
<br>
<br>

<hr>
<hr>

en stocks and bonds? 


**Have fun**
